---
title: "Spatial clustering to define spatial scales and synthetic habitat covariates"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
# source("../functions/spatial_clustering.R")
source("../functions/data_processing.R")
library(ade4)
library(adegraphics)
library(sf)

covariates <- readRDS("covariates.rds")
```

# Camera-level covariates
Let's establish clusters of similar cameras in terms of habitat in order to build a synthetic habitat covariate.

```{r, include = FALSE}
sort(covariates$dist_road)
unique(covariates$habitat_mucina)
sort(covariates$avg_trees)
sort(covariates$dist_water)
```

We perform a multivariate analysis on the data:

```{r}
# Prepare data for mv analysis
cov.mva <- covariates %>% select(-code_loc) %>%
        st_drop_geometry() %>% remove_rownames() %>%  column_to_rownames("site_ID")
cov.mva

hillsmith <- dudi.hillsmith(cov.mva, scannf = FALSE, nf = 3)

scatter(hillsmith, plabels.optim = TRUE)
```

Then, we perform hierarchical clustering for habitat covariates at camera level to define the synthetic habitat covariate classes.

```{r}
dist <- dist.dudi(hillsmith)

fit <- hclust(dist, method = "complete")
plot(fit)

threshold <- 5.5

covcam <- cutree(fit, h = threshold)

f <- covcam[match(rownames(hillsmith$li), names(covcam))]
fact <- as.factor(f)

s.class(hillsmith$li, fac = fact,
        ellipse = 0, starSize = 0, chullSize = 1,
        col = TRUE)
```

Spatial visialisation of the habitat covariates associated with cameras:

```{r, echo = FALSE}
cluplot <- as.data.frame(covcam) %>% rownames_to_column("site_ID")
plot <- covariates %>% left_join(cluplot, by = 'site_ID')

ggplot() + geom_sf(data = plot, aes(col = as.factor(covcam)), show.legend = FALSE) + 
        theme_bw() + ggtitle("Habitat covariates")
```

# Not constrained clustering
Now we build camera clusters based on distance to get different inference scales.

## Sites clustering (get different scales)

```{r}
clust <- covariates %>% dplyr::select(site_ID)

#distance matrix
distmat <- st_distance(clust)
rownames(distmat) <- clust$site_ID
colnames(distmat) <- clust$site_ID

#clustering
fit <- hclust(as.dist(distmat), method = "complete")
plot(fit)
```

```{r}
# Threshold
threshold.coarse <- 6000
threshold.fine <- 2500

# Tree cutting
clusters.coarse <- cutree(fit, h = threshold.coarse)

clusters.fine <- cutree(fit, h = threshold.fine)

# Plotting
tab <- data.frame(clusters.coarse) %>% rownames_to_column("site_ID")
tab2 <- data.frame(clusters.fine) %>% rownames_to_column("site_ID")
tab <- tab %>% left_join(tab2, by = "site_ID")

clust <- clust %>% left_join(tab, by = "site_ID")
```

Result:

```{r, echo = FALSE}
ggplot() + geom_sf(data = clust, aes(col = as.factor(clusters.fine))) + ggtitle("Fine clusters")
ggplot() + geom_sf(data = clust, aes(col = as.factor(clusters.coarse))) + ggtitle("Coarse clusters")
```

## Sites covariates
Replace the original camera-level covariate by mean/new factor level.
```{r}
cov <- covariates %>% left_join(tab, by = "site_ID")

cov.fine <- cov %>% st_drop_geometry() %>% group_by(clusters.fine) %>%
        summarise(mean_road = mean(dist_road),
                  mean_trees = mean(avg_trees),
                  mean_water =  mean(dist_water),
                  habitat_mucina = as.factor(paste(unique(habitat_mucina), collapse = ".")))  %>%
        mutate(clusters.fine = paste0("SAM", clusters.fine)) %>%
        column_to_rownames("clusters.fine")
cov.fine

cov.coarse <- cov %>% st_drop_geometry() %>% group_by(clusters.coarse) %>%
        summarise(mean_road = mean(dist_road),
                  mean_trees = mean(avg_trees),
                  mean_water =  mean(dist_water),
                  habitat_mucina = as.factor(paste(unique(habitat_mucina), collapse = ".")))  %>%
        mutate(clusters.coarse = paste0("SAM", clusters.coarse)) %>%
        column_to_rownames("clusters.coarse")
cov.coarse
```

```{r}
hillsmith.coarse <- dudi.hillsmith(cov.coarse, nf = 2, scannf = FALSE)
dist.cov.coarse <- dist.dudi(hillsmith.coarse)

fit.coarse <- hclust(dist.cov.coarse, method = "complete")
plot(fit.coarse)

threshold.coarse <- 3.8
cov.coarse <- cutree(fit.coarse, h = threshold.coarse)
```

```{r}
hillsmith.fine <- dudi.hillsmith(cov.fine, nf = 3, scannf = FALSE)
dist.cov.fine <- dist.dudi(hillsmith.fine)

fit.fine <- hclust(dist.cov.fine, method = "complete")
plot(fit.fine)

threshold.fine <- 5
cov.fine <- cutree(fit.fine, h = threshold.fine)
```

Final cluster covariates:
```{r}
cov.coarse
cov.fine
```

